---
title: "pjt_bank_marketing"
author: "Eunyoung Kwak, Cindy Miao, Haiying Lin, Susie Liang"
date: "2024-05-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, label='Libraries', setup, include=FALSE}
library(readxl)
library(data.table)
library(tidyverse)
library(randomForest) # Random Forests
library(e1071) # SVM
library(gridExtra)
library(cluster) # cluster
library(doParallel)
library(factoextra)
library(rpart)
library(rpart.plot)
library(ggplot2)
library(caret) #lasso split data
library(glmnet) # lasso model
```


# 1. Data Preparation

* Banking Dataset - Marketing Targets
  + https://www.kaggle.com/datasets/prakharrathi25/banking-dataset-marketing-targets

```{r Data Structure}
bank <- read_excel("bank.xlsx") # input data
str(bank)
```

## 7 Numarical Variables

```{r Numerical Summaries}
summary(bank[sapply(bank, is.numeric)])
```

## 10 Categorical Variables (including the target variable y)

```{r Categorical Variables}
sapply(bank[sapply(bank, is.character)], unique)
```

```{r Factors}
bank[sapply(bank, is.character)] <- lapply(bank[sapply(bank, is.character)], as.factor)
bank$month <- factor(bank$month, levels = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec")) 
bank$poutcome <- factor(bank$poutcome, levels = c("success", "failure", "unknown", "other"))
# str(bank)
summary(bank[sapply(bank, is.factor)])
```

# 2. Exploratory Data Analysis

## Target Variable y Term deposit subscribed?

```{r Y Variable}

# y_count <- table(bank$y)
# y_label <- paste0(toupper(names(y_count))," - ",y_count," (",round(y_count/sum(y_count)*100,1),"%)")
# 
# pie(y_count, labels=y_label, col=c("#EE4E4E","#97BE5A"), main="Term deposit subscribed")

tibble(count = as.integer(table(bank$y)),
       label = paste0(toupper(names(table(bank$y)))," - ",table(bank$y)," (",round(table(bank$y)/sum(table(bank$y))*100,1),"%)")) %>%
  ggplot(aes(x = "", y = count, fill = label)) +
  geom_bar(stat = "identity", width = 1, show.legend = FALSE) +
  coord_polar(theta = "y") +
  theme_void() +
  labs(title = "Term Deposit Subscription") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5))

```

## Predictor variables: Numerical

* Most of them are right-skewed and have a lot of outliers.

```{r Numerical-Hist}
colnm1 <- colnames(bank[sapply(bank, is.numeric)])

plot_list <- list()

for (i in  1:length(colnm1)) {
  plot_list[[i]] <- ggplot(bank) +
    geom_histogram(aes(x=!!sym(colnm1[i])), fill="orange", show.legend = FALSE, bins=30) +
    ggtitle(colnm1[i]) + 
    theme_bw()
}

grid.arrange(plot_list[[1]], plot_list[[2]], plot_list[[3]], plot_list[[4]], ncol=2)
grid.arrange(plot_list[[5]], plot_list[[6]], plot_list[[7]], ncol=2)

```

```{r Numerical-Box}
colnm1 <- colnames(bank[sapply(bank, is.numeric)])

plot_list <- list()

for (i in  1:length(colnm1)) {
  plot_list[[i]] <- ggplot(bank) +
    geom_boxplot(aes(x=!!sym(colnm1[i]))) +
    ggtitle(colnm1[i]) + coord_flip() +
    theme_bw()
}

grid.arrange(plot_list[[1]], plot_list[[2]], plot_list[[3]], plot_list[[4]], ncol=4)
grid.arrange(plot_list[[5]], plot_list[[6]], plot_list[[7]], ncol=3)
```

## Predictor variables: Categorical

* contact and poutcome columns have relatively large portion of "unknown" values.

```{r Categorical - Bar}
desc <- c("Type of Job", "Marital status", "Education level", "Has credit in default?",
          "Has a housing loan?", "Has a personal loan?", "Contact communication type",
          "Last contact month of year", "Previous campaign results", "Term deposit subscribed?")
colnm2 <- colnames(bank[sapply(bank, is.factor)] %>% select(-y))

plot_list <- list()

for (i in 1:length(colnm2)) {
  plot_list[[i]] <- ggplot(data = bank) + 
          geom_bar(aes(x = !!sym(colnm2[i]), fill=!!sym(colnm2[i])), show.legend = FALSE) + 
          coord_flip() +
          ggtitle(paste0(colnm2[i],": ",desc[i])) + xlab("") + theme_bw()
}

plot_list[[1]]
grid.arrange(plot_list[[2]], plot_list[[3]], ncol=1)
grid.arrange(plot_list[[4]], plot_list[[5]], ncol=1)
grid.arrange(plot_list[[6]], plot_list[[7]], ncol=1)
grid.arrange(plot_list[[8]], plot_list[[9]], ncol=1)

```

## Relatioships

### Correlation

* pdays - poutcome(unknown) : -0.87
* pdays - poutcome(failure) :  0.70
* education(tertiary) - job(management) : 0.60
* poutcome(unknown) - previous : -0.53

```{r Highest Cor}
matrix.bank <- model.matrix(~., data=bank)
matrix.bank <- matrix.bank[,-1] ## remove intercept

data.frame(which(abs(cor(matrix.bank)) > 0.5 & cor(matrix.bank) != 1, arr.ind = TRUE)) %>%
  mutate(rowname=rownames(cor(matrix.bank))[row],
         colname=colnames(cor(matrix.bank))[col],
         corr = round(cor(matrix.bank)[cbind(row, col)],2)) %>%
  select(3:5) %>%
  arrange(desc(abs(corr)))
```

*pdays - poutcome* 
  + If the outcome of the last campaign was unknown, the bank re-contacts the customer as soon as possible. (correlation: -0.87)
  + If successful, the bank re-contacts the customer relatively quickly because this campaign is likely to be successful as well.
  + If the campaign was unsuccessful (failure or other), the bank reaches out to the customer at a longer interval because this campaign is likely to be unsuccessful as well. (correlation: 0.70)

```{r pdays - poutcome}

ggplot(bank, aes(x=reorder(poutcome, pdays, FUN=median), y=pdays)) +
  geom_boxplot() +
  labs(title="Days since last contact in previous campaign vs. previous campaign outcome", x="previous campaign outcome") +
  theme_bw()

```

*education(tertiary) - job(management): 0.60*

```{r education(tertiary) - job(management)}

bank %>% select(job, education) %>%
  group_by(job, education) %>% summarise(cnt=n(), .groups="drop") %>%
  group_by(job) %>% mutate(Percent = round(cnt/sum(cnt)*100,1)) %>%

  ggplot(aes(x=job, y=Percent, fill=education)) +
  geom_bar(stat="identity") +
  geom_text(aes(label=paste0(Percent,"%")), position = position_stack(vjust = 0.5)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

*poutcome(unknown) - previous: -0.53*
Previous contacts for this client before this campaign vs. previous campaign outcome

```{r poutcome(unknown) - previous}

gg1 <- ggplot(bank, aes(x=reorder(poutcome, previous, FUN=median), y=previous)) +
  geom_boxplot() +
  labs(title="Previous vs. previous outcome", x="previous campaign outcome") +
  theme_bw()

gg2 <- gg1 + coord_cartesian(ylim=c(0,10))

grid.arrange(gg1, gg2, ncol=2)

```


### Relationships between Target Variable(y) and Predictor Variables

```{r Predictor with y Percent}

# Early Adult: 18-29 years old
# Young Adult: 30-44 years old
# Middle-Aged Adult: 45-64 years old
# Elderly/Senior: 65 years old and older

desc <- c("Type of Job", "Marital status", "Education level", "Has credit in default?",
          "Has a housing loan?", "Has a personal loan?", "Contact communication type",
          "Last contact month of year", "Previous campaign results", "Term deposit subscribed?")

colnm <- colnames(bank[sapply(bank, is.factor)] %>% select(-y))

for (i in  1:length(colnm)) {
  print(bank %>%
      select(y,age,!!sym(colnm[i])) %>% 
      mutate(AgeGroup=cut(bank$age, breaks = c(18, 30, 45, 64, 96),
                          labels = c("18-29", "30-44", "45-64", "65-95"),
                          right = FALSE)) %>%
      group_by(!!sym(colnm[i]), AgeGroup) %>%
      summarize(cnt=n(), .groups = "drop") %>%
      group_by(!!sym(colnm[i])) %>%
      mutate(Percent=round(cnt/sum(cnt)*100,1)) %>%
      ggplot(aes(x=!!sym(colnm[i]), y=Percent, fill=AgeGroup)) + 
      geom_bar(stat="identity") +
      geom_text(aes(label=paste0(Percent,"%")), position = position_stack(vjust = 0.5)) +
      theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
      labs(title=desc[i], x="", y=""))
}

```

#### Numerical variables vs. y

```{r Numerical vs. y - Density}

gg1 <- ggplot(bank, aes(x=age, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg2 <- ggplot(bank, aes(x=balance, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg3 <- ggplot(bank, aes(x=day, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg4 <- ggplot(bank, aes(x=duration, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg5 <- ggplot(bank, aes(x=campaign, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg6 <- ggplot(bank, aes(x=pdays, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()
gg7 <- ggplot(bank, aes(x=previous, fill=y)) + geom_density(alpha=0.3) + labs(fill="subscribed") + theme_bw()

grid.arrange(gg1, gg2, gg3, gg4, ncol=2)
grid.arrange(gg5, gg6, gg7, ncol=2)

```

```{r Numerical vs. y (Density scaled)}
ggplot(bank, aes(x=balance, fill=y)) + geom_density(alpha=0.3) + labs(fill="Subscribed") + 
  coord_cartesian(xlim=c(-2000, 10000)) + theme_bw()
ggplot(bank, aes(x=duration, fill=y)) + geom_density(alpha=0.3) + labs(fill="Subscribed") + 
  coord_cartesian(xlim=c(0,2000)) + theme_bw()
ggplot(bank, aes(x=campaign, fill=y)) + geom_density(alpha=0.3) + labs(fill="Subscribed") + 
  coord_cartesian(xlim=c(0,20)) + theme_bw()
ggplot(bank, aes(x=pdays, fill=y)) + geom_density(alpha=0.3) + labs(fill="Subscribed") + 
  coord_cartesian(xlim=c(0,250)) + theme_bw()
ggplot(bank, aes(x=previous, fill=y)) + geom_density(alpha=0.3) + labs(fill="Subscribed") + 
  coord_cartesian(xlim=c(0,15)) + theme_bw()
```

*pdays: Number of days that passed by after the client was last contacted from a previous campaign (numeric, -1 means client was not previously contacted)*

```{r pdays vs. y - Bar}
bank %>%
  select(y,pdays) %>% 
  mutate(pday=cut(bank$pdays, breaks = c(-1, 0, 50, 100, 150, 200, 250, 300, Inf),
                      labels = c("No Contact", "0-50", "50-100", "100-150", "150-200", "200-250", "250-300", "300+"),
                      right = FALSE)) %>%
  group_by(y, pday) %>%
  summarise(cnt=n(), .groups = "drop") %>%
  group_by(y) %>%
  mutate(Percent=round(cnt/sum(cnt)*100,1)) %>%
  ggplot(aes(x=pday, y=Percent, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  labs(title="Days since last contact in previous campaign", x="", y="% Subscribed", fill="Subscribed") +
  
  theme_bw()
```

```{r Numerical vs. y - Box}

gg1 <- bank %>% ggplot(aes(x=y, y=age)) + geom_boxplot() + theme_bw()
gg2 <- bank %>% ggplot(aes(x=y, y=balance)) + geom_boxplot() + theme_bw()
gg3 <- bank %>% ggplot(aes(x=y, y=day)) + geom_boxplot() + theme_bw()
gg4 <- bank %>% ggplot(aes(x=y, y=duration)) + geom_boxplot() + theme_bw()
gg5 <- bank %>% ggplot(aes(x=y, y=campaign)) + geom_boxplot() + theme_bw()
gg6 <- bank %>% ggplot(aes(x=y, y=pdays)) + geom_boxplot() + theme_bw()
gg7 <- bank %>% ggplot(aes(x=y, y=previous)) + geom_boxplot() + theme_bw()

grid.arrange(gg1, gg2, gg3, gg4, gg5, gg6, gg7, ncol=4)

```

```{r Categorical vs. y Percent - Bar}

desc <- c("Type of Job", "Marital status", "Education level", "Has credit in default?", "Has a housing loan?",
          "Has a personal loan?", "Contact communication type", "Last contact month of year", "Previous campaign results")
colnm2 <- colnames(bank[sapply(bank, is.factor)] %>% select(-y))

for (i in 1:length(colnm2)) {
  print(bank %>% 
    group_by(y,!!sym(colnm2[i])) %>% summarize(cnt=n(), .groups="drop") %>% 
    group_by(y) %>% mutate(Percent = cnt/sum(cnt)*100) %>%
    ggplot(aes(x = reorder(!!sym(colnm2[i]), Percent, FUN=max), y = Percent, fill = y)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_text(aes(label=paste(round(Percent,2),"%")), position = position_dodge(width = 0.9), hjust = -0.1, size=3.5) +
    labs(title = desc[i],
         x = "",
         y = "% Subscription",
         fill = "Subscribed") +
    coord_flip() +
    theme_bw())
}

```

# 3. Customer Segmentation --- K means

```{r Data Preparation}
# Scale only numeric variables
numeric_vars <- bank %>% select(where(is.numeric))
scaled_numeric_vars <- as.data.frame(scale(numeric_vars))

# Combine the scaled numeric variables with the categorical variables
categorical_vars <- bank %>% select(where(is.factor))
scaled_bank <- cbind(scaled_numeric_vars, categorical_vars)

# Convert only categorical variables to dummy variables
scaled_bank_dummies <- scaled_bank %>%
  mutate_if(is.factor, as.character) %>%
  mutate_if(is.character, as.factor) %>%
  model.matrix(~ . -1, data = .) %>%
  as.data.table()
```


## Elbow Test
```{r Elbow Test, warning=FALSE}
# # Elbow Test
# set.seed(208) # For reproducibility
# 
# # Register parallel backend
# cl <- makeCluster(detectCores() - 1) # Use one less core than available
# registerDoParallel(cl)
# 
# # Sample a subset of the data
# sampled_data <- scaled_bank_dummies[sample(1:nrow(scaled_bank_dummies), size = 1000), ]
# 
# # Run fviz_nbclust with parallel processing 
# fviz_nbclust(sampled_data, kmeans, method = "wss", k.max = 10)
# 
# # Stop the parallel backend
# stopCluster(cl)
# registerDoSEQ()

set.seed(208)

wss <- function(k) {
  kmeans(scaled_bank_dummies, k)$tot.withinss
}
k.values <- 1:15
wss.values <- sapply(k.values, wss)

plot(k.values, wss.values, type="b", pch = 19, frame = FALSE,
     xlab="Number of clusters K",
     ylab="Total within-clusters sum of squares")


```

**Based on the elbow plot shown, the optimal number of clusters to be 3.**


## K-Means Cluster Plot
```{r K-Means Clusting}
# Apply k-means clustering

# optimal 
set.seed(208)
StrattonCluster_3k <- kmeans(scaled_bank_dummies, 3)
set.seed(208)
StrattonCluster_4k <- kmeans(scaled_bank_dummies, 4)
set.seed(208)
StrattonCluster_5k <- kmeans(scaled_bank_dummies, 5)
set.seed(208)
StrattonCluster_6k <- kmeans(scaled_bank_dummies, 6)
```

*I changed the x, y scaling to exclude the most skewed outliers.*
```{r 3 Cluster Plot}
set.seed(208)
# # Visualize the clusters k = 3
fviz_cluster(StrattonCluster_3k, data = scaled_bank_dummies, 
              ellipse.type = "norm", geom = "point", pointsize = 1) +
   ggtitle("Cluster Visualization with Custom Scales") +
   theme_minimal()
 
# # Visualize the clusters k = 3 without outlier
 fviz_cluster(StrattonCluster_3k, data = scaled_bank_dummies, 
              ellipse.type = "norm", geom = "point", pointsize = 1) +
   xlim(-5, 15) + # Adjust x-axis limits
   ylim(-10, 10) + # Adjust y-axis limits
   ggtitle("Cluster Visualization with Custom Scales") +
   theme_minimal()
```

```{r 456 Cluster K-Means}
set.seed(208)
# Visualize the clusters k = 4
Cluster4 <- fviz_cluster(StrattonCluster_4k, data = scaled_bank_dummies, 
             ellipse.type = "norm", geom = "point", pointsize = 1) +
  xlim(-5, 15) + # Adjust x-axis limits
  ylim(-10, 10) + # Adjust y-axis limits
  ggtitle("4 Clusters") +
  theme_minimal()

set.seed(208)
# Visualize the clusters k = 5
Cluster5 <- fviz_cluster(StrattonCluster_5k, data = scaled_bank_dummies, 
             ellipse.type = "norm", geom = "point", pointsize = 1) +
  xlim(-5, 15) + # Adjust x-axis limits
  ylim(-10, 10) + # Adjust y-axis limits
  ggtitle("5 Clusters") +
  theme_minimal()

set.seed(208)
# Visualize the clusters k = 6
Cluster6 <- fviz_cluster(StrattonCluster_6k, data = scaled_bank_dummies, 
             ellipse.type = "norm", geom = "point", pointsize = 1) +
  xlim(-5, 15) + # Adjust x-axis limits
  ylim(-10, 10) + # Adjust y-axis limits
  ggtitle("6 Clusters") +
  theme_minimal()

#grid.arrange(Cluster4, Cluster5, Cluster6, ncol=3, nrow = 1)
Cluster4
Cluster5
Cluster6
```
**Based on the cluster plot, it seems that customers should be divided into 3 clusters.**

##  Cluster Size Plots
```{r 4 Cluster Size Plot}
# Cluster size of each cluster
sizes4k <- data.frame(Size = StrattonCluster_4k[["size"]], 
                      Cluster = c("Cluster1", "Cluster2", "Cluster3", "Cluster4"))

size4 <- ggplot(sizes4k, aes(factor(Cluster), Size)) + 
  geom_col(aes(fill = Cluster)) + 
  xlab("Cluster") + 
  ylab("Size") + 
  geom_text(aes(label=Size), vjust = -0.2) + 
  ggtitle("Cluster Size 4-Clusters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Paired")
```

```{r 5 Cluster Size Plot}
sizes5k <- data.frame(Size = StrattonCluster_5k[["size"]], 
                      Cluster = c("Cluster1", "Cluster2", "Cluster3", "Cluster4", "Cluster5"))

size5 <- ggplot(sizes5k, aes(factor(Cluster), Size)) + 
  geom_col(aes(fill = Cluster)) + 
  xlab("Cluster") + 
  ylab("Size") + 
  geom_text(aes(label=Size), vjust = -0.2) + 
  ggtitle("Cluster Size 5-Clusters") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Paired")
```

```{r 6 Cluster Size Plot, echo=FALSE, fig.height=8, fig.width=11}
sizes6k <- data.frame(Size = StrattonCluster_6k[["size"]], 
                      Cluster = c("Cluster1", "Cluster2", "Cluster3", "Cluster4", "Cluster5", "Cluster6"))

size6 <- ggplot(sizes6k, aes(factor(Cluster), Size)) + 
  geom_col(aes(fill = Cluster)) + 
  xlab("Cluster") + 
  ylab("Size") + 
  geom_text(aes(label=Size), vjust = -0.2) + 
  ggtitle("Cluster Size 6-Clusters") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Paired")

grid.arrange(size4, size5, size6, ncol=3, nrow = 1)
```

```{r Add Clusters into Original Data}
# Include Clusters into original data
bank <- bank %>% 
  mutate(k3Cluster = as.factor(StrattonCluster_3k[["cluster"]]),
         k4Cluster = as.factor(StrattonCluster_4k[["cluster"]]),
         k5Cluster = as.factor(StrattonCluster_5k[["cluster"]]),
         k6Cluster = as.factor(StrattonCluster_6k[["cluster"]]))
```

## Subscription Rate Per Cluster Plots
```{r 3 Clusters}
bank %>%
  group_by(k3Cluster, y) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(x=k3Cluster, y=cnt, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=cnt), position = position_dodge(width = 0.9), vjust=-0.2)

bank %>%
  group_by(k3Cluster, y) %>%
  summarise(cnt = n()) %>%
  ungroup() %>%
  group_by(k3Cluster) %>%
  mutate(total = sum(cnt),
         percentage = cnt / total * 100) %>%
  ggplot(aes(x=k3Cluster, y=percentage, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=paste0(round(percentage,1), "%")), position = position_dodge(width = 0.9), vjust=-0.2)
```

```{r 456 Clusters Subscription Rate, fig.height=8, fig.width=11}
# 4 Clusters
ycount4 <- bank %>%
  group_by(k4Cluster, y) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(x=k4Cluster, y=cnt, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=cnt), position = position_dodge(width = 0.9), vjust=-0.2)

y4 <- bank %>%
  group_by(k4Cluster, y) %>%
  summarise(cnt = n()) %>%
  ungroup() %>%
  group_by(k4Cluster) %>%
  mutate(total = sum(cnt),
         percentage = cnt / total * 100) %>%
  ggplot(aes(x=k4Cluster, y=percentage, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  ggtitle("Subsciption Rate 4 Clusters") +
  geom_text(aes(label=paste0(round(percentage,1), "%")), position = position_dodge(width = 0.9), vjust=-0.2)

# 5 Clusters
ycount5 <- bank %>%
  group_by(k5Cluster, y) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(x=k5Cluster, y=cnt, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=cnt), position = position_dodge(width = 0.9), vjust=-0.2)

y5 <- bank %>%
  group_by(k5Cluster, y) %>%
  summarise(cnt = n()) %>%
  ungroup() %>%
  group_by(k5Cluster) %>%
  mutate(total = sum(cnt),
         percentage = cnt / total * 100) %>%
  ggplot(aes(x=k5Cluster, y=percentage, fill=y)) +
  geom_bar(stat="identity", position="dodge") + 
  ggtitle("Subsciption Rate 5 Clusters") +
  geom_text(aes(label=paste0(round(percentage,1), "%")), position = position_dodge(width = 0.9), vjust=-0.2)

# 6 Clusters
ycount6 <- bank %>%
  group_by(k6Cluster, y) %>%
  summarise(cnt = n()) %>%
  ggplot(aes(x=k6Cluster, y=cnt, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  geom_text(aes(label=cnt), position = position_dodge(width = 0.9), vjust=-0.2)

y6 <- bank %>%
  group_by(k6Cluster, y) %>%
  summarise(cnt = n()) %>%
  ungroup() %>%
  group_by(k6Cluster) %>%
  mutate(total = sum(cnt),
         percentage = cnt / total * 100) %>%
  ggplot(aes(x=k6Cluster, y=percentage, fill=y)) +
  geom_bar(stat="identity", position="dodge") +
  ggtitle("Subsciption Rate 6 Clusters") +
  geom_text(aes(label=paste0(round(percentage,1), "%")), position = position_dodge(width = 0.9), vjust=-0.2)

grid.arrange(ycount4, ycount5, ycount6, ncol=3, nrow = 1)

grid.arrange(y4, y5, y6, ncol=3, nrow = 1) 
```

## Cluster Mean Plots and Cluster Profile
```{r 456 Cluster Means, fig.height=10, fig.width=13}
# 4 Clusters
# Build Mean per Cluster DataFrame
scaled_bank_dummies$k4Cluster = StrattonCluster_4k[["cluster"]]

summarystats.percluster_4k = scaled_bank_dummies %>% group_by(k4Cluster) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

var <- summarystats.percluster_4k %>% 
  gather(key = "variable", value = "value", -k4Cluster)

ggplot(var, aes(factor(k4Cluster), value)) +
  geom_bar(stat = "identity", aes(fill = variable), show.legend = FALSE)  + 
  xlab("Clusters") +
  geom_text(aes(label = round(value, digits = 3)), size = 3.5, colour = "white", 
            position = position_stack(vjust = 0.5)) +
  ggtitle("Cluster Means of Each Variable 4 Clusters") +
  facet_wrap(~variable, ncol=7, scales = "free_y") +
  theme(strip.text = element_text(size = 10, face = "bold"),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())

# 5 Clusters
# Build Mean per Cluster DataFrame
scaled_bank_dummies$k5Cluster = StrattonCluster_5k[["cluster"]]

summarystats.percluster_5k = scaled_bank_dummies %>% group_by(k5Cluster) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

var <- summarystats.percluster_5k %>% 
  gather(key = "variable", value = "value", -k4Cluster, -k5Cluster)

ggplot(var, aes(factor(k5Cluster), value)) +
  geom_bar(stat = "identity", aes(fill = variable), show.legend = FALSE)  + 
  xlab("Clusters") +
  geom_text(aes(label = round(value, digits = 3)), size = 3.5, colour = "white", 
            position = position_stack(vjust = 0.5)) +
  ggtitle("Cluster Means of Each Variable 5 Clusters") +
  facet_wrap(~variable, ncol=7, scales = "free_y") +
  theme(strip.text = element_text(size = 10, face = "bold"),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())

# 6 Clusters
# Build Mean per Cluster DataFrame
scaled_bank_dummies$k6Cluster = StrattonCluster_6k[["cluster"]]

summarystats.percluster_6k = scaled_bank_dummies %>% group_by(k6Cluster) %>% 
  summarise_if(is.numeric, mean, na.rm = TRUE)

var <- summarystats.percluster_6k %>% 
  gather(key = "variable", value = "value", -k4Cluster, -k5Cluster, -k6Cluster)

ggplot(var, aes(factor(k6Cluster), value)) +
  geom_bar(stat = "identity", aes(fill = variable), show.legend = FALSE)  + 
  xlab("Clusters") +
  geom_text(aes(label = round(value, digits = 3)), size = 3.5, colour = "white", 
            position = position_stack(vjust = 0.5)) +
  ggtitle("Cluster Means of Each Variable 6 Clusters") +
  facet_wrap(~variable, ncol=7, scales = "free_y") +
  theme(strip.text = element_text(size = 10, face = "bold"),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())
```

## Cluster 2 Logistic
```{r Full Model Logistic}
bank$y <- ifelse(bank$y=="yes", 1, 0) %>% as.factor(.) # 1=yes,0=no 

# full model
bank1 <- bank %>% 
  select(-k3Cluster, -k5Cluster, -k6Cluster) %>%
  filter(k4Cluster == 2) %>%
  select(-k4Cluster)

logistic_cluster2 <- glm(y ~ ., data = bank1, family = "binomial")

summary(logistic_cluster2)
```

```{r Check Collinearity}
library(performance)
check_collinearity(logistic_cluster2)
```

```{r Selected Model}
# Remove month and poutcome
cluster2 <- bank1 %>%
  select(-month, -poutcome)

# Train the logistic regression model using Cluster 2
logistic_cluster2 <- glm(y ~ ., data = cluster2, family = "binomial")

summary(logistic_cluster2)
```

```{r MultiCollinearity}
library(performance)
check_collinearity(logistic_cluster2)
```

```{r Odds}
# Odds Ratio
exp(cbind(OR=coef(logistic_cluster2), confint(logistic_cluster2)))
```


# 4. Lasso Variable Selection

## Split data into training dataset and testing dataset
```{r Data Partition, warning=FALSE}
# remove clusters
bank <- bank %>% select(-k3Cluster, -k4Cluster, -k5Cluster, -k6Cluster)

# Split the data into two sets based on the value of y
bank_y1 <- bank %>% filter(y == 1)
bank_y0 <- bank %>% filter(y == 0)

# Function to split data into train (80%) and test (20%) using caret
split_data <- function(data, train_ratio = 0.8) {
  set.seed(208) # For reproducibility
  train_indices <- createDataPartition(data$y, p = train_ratio, list = FALSE)
  
  train_set <- data[train_indices, ]
  test_set <- data[-train_indices, ]
  
  list(train = train_set, test = test_set)
}

# Split each dataset into train and test sets
split_y1 <- split_data(bank_y1)
split_y0 <- split_data(bank_y0)

# Extract train and test sets
train_y1 <- split_y1$train
test_y1 <- split_y1$test
train_y0 <- split_y0$train
test_y0 <- split_y0$test

# Combine the train and test sets
train_data <- bind_rows(train_y1, train_y0)
test_data <- bind_rows(test_y1, test_y0)
```

```{r Lambda vs. cv Error}
x <- model.matrix(y ~ ., train_data)[, -17]
y <- train_data$y

cv.lasso <- cv.glmnet(x, y, alpha = 1, family = "binomial")
plot(cv.lasso)
```

```{r Lasso Final Model}
# exact value of lambda
cv.lasso$lambda.min

# Final model with lambda.min
lasso_model <- glmnet(x, y, alpha = 1, family = "binomial",
                      lambda = cv.lasso$lambda.min)
lasso_model
```
```{r}
coef(lasso_model, lasso_model$lambda.min)
```

```{r Lasso Prediction}
# Make prediction on test data
x.test <- model.matrix(y ~ ., test_data)[, -17]
y.test <- test_data$y

pred_prob <- lasso_model %>% predict(newx = x.test)
pred_class <- ifelse(pred_prob > 0.8, 1, 0)

# Evaluate the model
confusionMatrix(as.factor(pred_class), as.factor(y.test), mode = "everything", positive = "1")
```


# 5. Predictive Modeling

## 5.1 Logistic Regression Analysis

### One Time Logistic
```{r Single Logistic}
# Train the logistic regression model using Cluster 2
logistic_model <- glm(y ~ ., data = bank, family = "binomial")

summary(logistic_model)
```

```{r Check MultiCollinearity}
# Check for Multicollinearity
library(performance)
check_collinearity(logistic_model)
```

```{r Odds Ratio}
# Odds Ratio
exp(cbind(OR=coef(logistic_model), confint(logistic_model)))
```

### 10-fold Cross Validation

```{r 10-fold CV Logistic}
# Set up 10-fold cross-validation
train_control <- trainControl(method = "cv", number = 10)

# Train the logistic regression model using 10-fold cross-validation
logistic_model_cv <- train(y ~ ., data = train_data, method = "glm", family = "binomial", trControl = train_control)

# Display the results
summary(logistic_model_cv)
```

```{r 10-fold cv Log Prediction}
# Make predictions on the test set
predictions <- predict(logistic_model_cv, newdata = test_data)
# Evaluate the model
confusionMatrix(predictions, test_data$y, mode = "everything", positive = "1")
```


## 5.2 SVM

```{r Tune SVM Polynomial}
svm.polynomial <- tune(svm, y~., data=bank, tunecontrol=tune.control(sampling="fix"),
                       kernel="polynomial", ranges=list(cost=2^seq(-3,5,1), degree=c(1,2,3)))
 
svm.polynomial$performances %>% arrange(error) %>% head(1)
 
svm.polynomial$performances %>%
    ggplot(aes(x=cost, y=error, color=as.factor(degree))) +
    geom_point() +
    geom_smooth(se=F, span=0.8) +
    labs(color="degree") +
    theme_bw()
```

```{r SVM Best Model}
# Extract the best model
best_svm_model <- svm.polynomial$best.model
```

```{r SVM Prediction}
# Make predictions on the test dataset
test_predictions <- predict(best_svm_model, newdata = test_data)
conf_matrix_test <- confusionMatrix(test_predictions, test_data$y,mode = "everything", positive="1")
conf_matrix_test
```

## 5.3 Decision tree

```{r Tree Model}
set.seed(208)
model.DT <- rpart(y~., data = bank, method = "class", cp = 0.001, minsplit = 400, maxdepth = 5)
rpart.plot(model.DT)
```

```{r Tree Prediction}
test_predictions <- predict(model.DT, newdata = test_data, type = "class")
conf_matrix_test <- confusionMatrix(test_predictions, test_data$y,mode = "everything",positive="1")
conf_matrix_test
```

## 5.4 Random Forests

```{r Random Forest}
set.seed(208)
results <- double(16)

for(m in 1:16) {
  model.RF <- randomForest(y~., data=bank, mtry=m, ntree=100)
  results[m] <- mean(model.RF$err.rate[,1])
}

tibble(m=1:16, error=results) %>%
  ggplot(aes(x=m, y=error)) +
  geom_point() +
  geom_smooth(se=F,span=0.4) +
  theme_bw()
```

```{r Random Forest Prediction}
which(results == min(results)) #7

model.RF <- randomForest(y~., data=train_data, mtry=10, ntree=100)

test_predictions <- predict(model.RF, newdata = test_data, type = "class")
conf_matrix_test <- confusionMatrix(test_predictions, test_data$y,mode = "everything",positive="1")

conf_matrix_test
```

```{r Variable Importance Plot}
varImpPlot(model.RF)
```

## 5.5 KNN
```{r KNN 10-fold CV Model}
set.seed(208)
knn_model_cv <- train(y ~ ., data = train_data, method = "knn", trControl = train_control, tuneLength = 10)
print(knn_model_cv)
```

```{r KNN Plot}
plot(knn_model_cv)
```

**This chart shows the k = 15 with accuracy approximately 88.69% for training dataset.**

```{r KNN Prediction}
prediction <- predict(knn_model_cv, newdata = test_data, mode = "everything", positive = "1")
cf <- confusionMatrix(prediction, test_data$y, mode = "everything", positive = "1")
cf
```

**With inital k = 15, the model correctly predict 88.55% target variable in test data set.**


